NanoMaLy

### 2015-11-12 paper/pldi16-submission/semantics.tex
define typeof

We can think of the evaluation of \eapp{f}{\ehole{}} as
computing a \emph{partial type} -- a type that may contain holes -- for
$f$. We can extract this type from the result of evaluation as follows.
%
\begin{defn}
\label{defn:partial-type}
  If \stepi{i}{\eapp{f}{\ehole{}}}{\emptysu}{e}{\su}, then the $i$th
  \emph{partial type} of f, written \ptype{i}{f},
  is \tfun{\typeof{\subst{\su}{\ehole{}}}}{\typeof{e}}.

  We will omit the subscript when we wish to refer to the final partial
  type. \ES{need to define typeof..}
\end{defn}
%
We also define a compatibility relation between types.
%
\begin{defn}
\label{defn:type-compat}
  A type $s$ is \emph{compatible} with a type $t$, written \tcompat{s}{t}, if
  $\exists \su.\ t = \subst{\su}{s} \lor s = \subst{\su}{t}$.
\end{defn}
%
Given these two definitions, we show that each evaluation step
refines the partial type of $f$.
%
\begin{lem}
\label{lem:refine-partial}
  For all $k$, \tcompat{\ptype{k}{f}}{\ptype{k+1}{f}}.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules. 
  %
  Note that all rules are type-preserving with the exception of when
  \forcesym is called on a hole, which case we instantiate the hole with
  a concrete value.
  %
  But \hastype{\ehole{}}{\thole{}}, which is compatible with any type.
\end{proof}
%
Furthermore, only a call to \forcesym can change the partial type of $f$.
%
\begin{lem}
\label{lem:force-inst}
  For all $k$, if $\ptype{k}{f} \neq \ptype{k+1}{f}$, then \forcesym must
  have been called at step $k+1$.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules. 
  % 
  If $\ptype{k}{f} \neq \ptype{k+1}{f}$ then one of the holes in $f$'s
  argument must have been instantiated with a concrete value at step
  $k+1$.
  %
  An examination of the rules shows that only place this happens is 
  in the first case of \forcesym.
\end{proof}
%
Finally, we show that any value that is not compatible with the $k$th
partial type of $f$ will cause $f$ to get stuck in \emph{at most} $k$
steps.
%
\begin{lem}
\label{lem:k-stuck}
  For any \hastype{v}{t} that is not compatible with
  the input of \ptype{k}{f}, \stepi{k}{\eapp{f}{v}}{\emptysu}{\stuck}{\su}.
\end{lem}
\begin{proof}
  By induction on $k$.

>>>>

We can think of the evaluation of \eapp{f}{\ehole{}} as computing a
\emph{partial type} -- a type that may contain holes -- for $f$.
%
We can extract this type from the result of evaluation as follows.
%
\begin{defn}
\label{defn:partial-type}
  If \stepi{i}{\eapp{f}{\ehole{}}}{\emptysu}{e}{\su}, then the $i$th
  \emph{partial type} of f, written \ptype{i}{f},
  is \tfun{\typeof{\subst{\su}{\ehole{}}}}{\typeof{e}}.

  We will omit the subscript when we wish to refer to the final partial
  type. \ES{need to define typeof..}
\end{defn}
%
The \typeof{} function is an approximation of the type of an expression.
\begin{defn}
\label{def:typeof}
  \[
  \begin{array}{lcll}
    \typeof{n}   & \defeq & \tint & \\
    \typeof{b}   & \defeq & \tbool & \\
    \typeof{\efun{x}{e}} & \defeq & \tfun{\thole{i}}{\typeof{e}}, & \quad \text{$i$ is fresh} \\
    \typeof{e} & \defeq & \thole{i}, & \quad \text{$i$ is fresh} \\
  \end{array}
  \]
\end{defn}
%
We also define a compatibility relation between types.
%
\begin{defn}
\label{defn:type-compat}
  A type $s$ is \emph{compatible} with a type $t$, written \tcompat{s}{t}, if
  $\exists \su.\ t = \subst{\su}{s} \lor s = \subst{\su}{t}$.
\end{defn}
%
Given these two definitions, we show that each evaluation step
refines the partial type of $f$.
%
\begin{lem}
\label{lem:refine-partial}
  For all $k$, \tcompat{\ptype{k}{f}}{\ptype{k+1}{f}}.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules. 
  %
  Note that all rules are type-preserving with the exception of when
  \forcesym is called on a hole, which case we instantiate the hole with
  a concrete value.
  %
  But \hastype{\ehole{}}{\thole{}}, which is compatible with any type.
\end{proof}
%
Furthermore, only a call to \forcesym can change the partial type of $f$.
%
\begin{lem}
\label{lem:force-inst}
  For all $k$, if $\ptype{k}{f} \neq \ptype{k+1}{f}$, then \forcesym must
  have been called at step $k+1$.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules. 
  % 
  If $\ptype{k}{f} \neq \ptype{k+1}{f}$ then one of the holes in $f$'s
  argument must have been instantiated with a concrete value at step
  $k+1$.
  %
  An examination of the rules shows that only place this happens is 
  in the first case of \forcesym.
\end{proof}
%
Finally, we show that any value that is not compatible with the $k$th
partial type of $f$ will cause $f$ to get stuck in \emph{at most} $k$
steps.
%
\begin{lem}
\label{lem:k-stuck}
  For any \hastype{v}{t} that is not compatible with
  the input of \ptype{k}{f}, \stepi{k}{\eapp{f}{v}}{\emptysu}{\stuck}{\su}.
\end{lem}
\begin{proof}
  By induction on $k$.
